---
title: "Exploring the relationships between a set of variables and miles per gallon for a collection of cars"
---

# Exploring the relationships between a set of variables and miles per gallon for a collection of cars

## Synopsis

Looking at a data set of a collection of cars, we are interested in exploring the relationship between a set of variables and miles per gallon (MPG). We are particularly interested in the following two questions:

* Is an automatic or manual transmission better for MPG
* Quantify the MPG difference between automatic and manual transmissions

## Exploring the data

Let's load dataset and look at the data primary features:

```{r}
library(datasets)

data(mtcars)

dim(mtcars)
head(mtcars)
summary(mtcars)
```

Dataset contains following variables:

1. mpg - Miles/(US) gallon
1. cyl - Number of cylinders
1. disp - Displacement (cu.in.)
1. hp - Gross horsepower
1. drat - Rear axle ratio
1. wt - Weight (lb/1000)
1. qsec - 1/4 mile time
1. vs - V/S
1. am - Transmission (0 = automatic, 1 = manual)
1. gear - Number of forward gears
1. carb - Number of carburetors

As we can see, there are bunch of variables that should be factors rather than numeric. Let's transofrm them appropriately:

```{r}
mtcars$cyl = factor(mtcars$cyl)
mtcars$vs = factor(mtcars$vs)
mtcars$am = factor(mtcars$am)
mtcars$gear = factor(mtcars$gear)
mtcars$carb = factor(mtcars$carb)

summary(mtcars)
```

Let's look at relationships between different variables:

```{r cache = TRUE}
library(GGally)
library(ggplot2)

ggpairs(mtcars, lower = list(continuous = "smooth"), params = c(method = "loess"))
```

## Building data model

Let's try to build data model based on all possible variables:

```{r}
fit1 <- lm(mpg ~ am, mtcars)
fit2 <- lm(mpg ~ am + cyl, mtcars)
fit3 <- lm(mpg ~ am + cyl + disp, mtcars)
fit4 <- lm(mpg ~ am + cyl + disp + hp, mtcars)
fit5 <- lm(mpg ~ am + cyl + disp + hp + drat, mtcars)
fit6 <- lm(mpg ~ am + cyl + disp + hp + drat + wt, mtcars)
fit7 <- lm(mpg ~ am + cyl + disp + hp + drat + wt + qsec, mtcars)
fit8 <- lm(mpg ~ am + cyl + disp + hp + drat + wt + qsec + vs, mtcars)
fit8 <- lm(mpg ~ am + cyl + disp + hp + drat + wt + qsec + vs + gear, mtcars)
fit9 <- lm(mpg ~ am + cyl + disp + hp + drat + wt + qsec + vs + gear + carb, mtcars)

anova(fit1, fit2, fit3, fit4, fit5, fit6, fit7, fit8, fit9)
```

As we can see, besides transmission information, adding number of cylinders and gross horsepower pass sygnificance test. Now, let's check if interactions between these values are significant:

```{r}
fit1 <- lm(mpg ~ am + cyl + hp, mtcars)
fit2 <- lm(mpg ~ am + cyl + hp + am*cyl, mtcars)
fit3 <- lm(mpg ~ am + cyl + hp + am*cyl + am*hp, mtcars)
fit4 <- lm(mpg ~ am + cyl + hp + am*cyl + am*hp + cyl*hp, mtcars)
fit5 <- lm(mpg ~ am + cyl + hp + am*cyl + am*hp + cyl*hp + am*cyl*hp, mtcars)
anova(fit1, fit2, fit3, fit4, fit5)
```

As we can see, interactions have no significance. So let's take original model:

```{r}
fit <- lm(mpg ~ am + cyl + hp, mtcars)
summary(fit)
```

As we can see, manual transmission brings more miles per gallon with mean value `r summary(fit)$coef[2,1]` and standard error `r summary(fit)$coef[2,2]`. So, this value is significant based on t-test, for >95% confidence interval.

Let's plot given model:

```{r}
ggplot(mtcars, aes(x = am, y = mpg, group = 1)) + 
geom_point(aes(color = cyl)) + 
geom_smooth(method = "lm", formula = y ~ x) +
labs(x = 'Transmission (0 = automatic, 1 = manual)') +
labs(y = 'Miles/(US) gallon') +
labs(title = 'Miles per gallon modelled by transmission type and number of cylinders')
```

```{r}
ggplot(mtcars, aes(x = am, y = mpg, group = 1)) + 
geom_point(aes(color = hp)) + 
geom_smooth(method = "lm", formula = y ~ x) +
labs(x = 'Transmission (0 = automatic, 1 = manual)') +
labs(y = 'Miles/(US) gallon') +
labs(title = 'Miles per gallon modelled by transmission type and gross horsepower')
```

And finally let's look at some diagnostics:

```{r}
par(mfrow = c(2, 2))
plot(fit)
```

As we can see, our model is acceptable and can be considered significant for investigations.

## Summary

So, final conclusion is that:
* we can build acceptable model of relation between miles per gallon value and transmission type, number of cylinders and gross horsepower
* manual transmission brings more miles per gallon with mean value `r summary(fit)$coef[2,1]` and standard error `r summary(fit)$coef[2,2]`
* manual transmission is significantly better than automatic for miles per gallon
