---
title: "Exploring the relationships between a set of variables and miles per gallon for a collection of cars"
---

# Exploring the relationships between a set of variables and miles per gallon for a collection of cars

## Synopsis

Looking at a data set of a collection of cars, we are interested in exploring the relationship between a set of variables and miles per gallon (MPG). We are particularly interested in the following two questions:

* Is an automatic or manual transmission better for MPG
* Quantify the MPG difference between automatic and manual transmissions

## Exploring the data

Let's load dataset and look at the data primary features:

```{r}
library(datasets)
data(mtcars)

dim(mtcars)
head(mtcars)
summary(mtcars)
```

As we can see, there are bunch of variables that should be factors rather than numeric. Let's transofrm them appropriately:

```{r}
mtcars$cyl = factor(mtcars$cyl)
mtcars$vs = factor(mtcars$vs)
mtcars$am = factor(mtcars$am)
mtcars$gear = factor(mtcars$gear)
mtcars$carb = factor(mtcars$carb)

summary(mtcars)
```

Let's look at relationships between different variables:

```{r}
plot(mtcars)
```

## Building data model

## Summary